---
title: "Model Comparison Project Example"
date: "Updated on `r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 10)
```

## Overview

This vignette allows you to walk through an example using `r4MAS` (R interface to the Metapopulation Assessment System) to run `MAS`. The example uses outputs from the operating model (OM) of [Age-structured stock assessment model comparison project](https://github.com/Bai-Li-NOAA/Age_Structured_Stock_Assessment_Model_Comparison) as inputs in `r4MAS`. It is an single-species, single-fleet, and single-survey example. The general workflow is:

1. Install packages
2. Load OM outputs
3. Load `r4MAS` module 
4. Specify inputs
5. Run `MAS` and check for convergence
6. Compare estimates from `MAS` with true values from the OM

## 1. Install packages
```{r setup_pkgs, eval= FALSE}
remotes::install_github("nmfs-fish-tools/r4MAS")
library(r4MAS)
library(Rcpp)
library(jsonlite)
library(nmfspalette)
```

If you receive errors related to C++ when loading r4MAS, please see the installation instructions from the [main page](https://github.com/nmfs-fish-tools/r4MAS#installation-instructions).

## 2. Load OM outputs
```{r}
data_path <- system.file("extdata", package = "r4MAS")
om_path <- file.path(data_path, "externalom_example")
load(file.path(om_path, "singlespecies.RData"))
```

The OM includes `om_input`, `om_output`, and `em_input` data lists.
```{r}
names(om_input)
```

```{r}
names(om_output)
```

```{r}
names(em_input)
```

## 3. Load `r4MAS` module 

Option 1:
```{r}
r4mas <- Rcpp::Module("rmas", PACKAGE = "r4MAS")
```

If you receive errors when using option 1, please use option 2:
```{r}
# Find the path of dynamically-loaded file with extension .so on Linux, .dylib on OS X or .dll on Windows
libs_path <- system.file("libs", package = "r4MAS")
dll_name <- paste("r4MAS", .Platform$dynlib.ext, sep = "")
if (.Platform$OS.type == "windows") {
  dll_path <- file.path(libs_path, .Platform$r_arch, dll_name)
} else {
  dll_path <- file.path(libs_path, dll_name)
}
r4mas <- Rcpp::Module("rmas", dyn.load(dll_path))
```
## 4. Specify inputs

### General settings in the inputs

- Population
    + Year, season, age, and area
    + Recruitment, growth, maturity parameters
    + Natural mortality
    + Movement 
    + Initial deviations
    + Sex ratio
- Fleet 
    + Index
    + Age composition
    + Likelihood components
    + Selectivity
    + Fishing mortality
- Survey
    + Index
    + Age composition
    + Likelihood components
    + Selectivity
    + Catchability
- Build the `MAS` model

### General settings of year, season, age, and area
```{r}
nyears <- om_input$nyr
nseasons <- 1
nages <- om_input$nages
ages <- om_input$ages
area1 <- new(r4mas$Area)
area1$name <- "area1"
```

### Recruitment settings
```{r}
recruitment <- new(r4mas$BevertonHoltRecruitment)
recruitment$R0$value <- om_input$R0 / 1000
recruitment$R0$estimated <- TRUE
recruitment$R0$phase <- 1
recruitment$h$value <- om_input$h
recruitment$h$estimated <- FALSE
recruitment$h$phase <- 3
recruitment$h$min <- 0.2001
recruitment$h$max <- 1.0
recruitment$sigma_r$value <- om_input$logR_sd
recruitment$sigma_r$estimated <- FALSE
recruitment$sigma_r$min <- 0
recruitment$sigma_r$max <- 1.0
recruitment$sigma_r$phase <- 2
recruitment$estimate_deviations <- TRUE
recruitment$constrained_deviations <- TRUE
recruitment$deviations_min <- -15.0
recruitment$deviations_max <- 15.0
recruitment$deviation_phase <- 2
recruitment$SetDeviations(om_input$logR.resid)
recruitment$use_bias_correction <- FALSE
```

### Growth settings

In this example, we directly use weight-at-age data for the assessment. 
```{r}
growth <- new(r4mas$VonBertalanffyModified)
empirical_weight <- rep(om_input$W.kg, times = om_input$nyr)
survey_empirical_weight <- replicate(nages * nyears, 1.0)
growth$SetUndifferentiatedCatchWeight(empirical_weight)
growth$SetUndifferentiatedWeightAtSeasonStart(empirical_weight)
growth$SetUndifferentiatedWeightAtSpawning(empirical_weight)
growth$SetUndifferentiatedSurveyWeight(survey_empirical_weight)
```
If you want to model length-at-age following a von Bertalanffy growth model, you can define the parameters related to growth.
```{r}
growth$a_min$value <- min(om_input$ages)
growth$a_max$value <- max(om_input$ages)
growth$c$value <- 0.3
growth$lmin$value <- 5
growth$lmax$value <- 50
growth$alpha_f$value <- om_input$a.lw
growth$alpha_m$value <- om_input$a.lw
growth$beta_f$value <- om_input$b.lw
growth$beta_m$value <- om_input$b.lw
```

### Maturity settings
```{r}
maturity <- new(r4mas$Maturity)
maturity$values <- om_input$mat.age * 0.5
```

### Natural mortality settings
```{r}
natural_mortality <- new(r4mas$NaturalMortality)
natural_mortality$SetValues(om_input$M.age)
```

### Movement settings
```{r}
# Only 1 area in this model
movement <- new(r4mas$Movement)
movement$connectivity_females <- c(0.0)
movement$connectivity_males <- c(0.0)
movement$connectivity_recruits <- c(0.0)
```

### Initial deviation settings
```{r}
initial_deviations <- new(r4mas$InitialDeviations)
initial_deviations$values <- rep(0.0, times = om_input$nages)
initial_deviations$estimate <- TRUE
initial_deviations$phase <- 2
```

### Creat the population
```{r}
population <- new(r4mas$Population)
for (y in 1:(nyears))
{
  population$AddMovement(movement$id, y)
}
population$AddNaturalMortality(natural_mortality$id, area1$id, "undifferentiated")
population$AddMaturity(maturity$id, area1$id, "undifferentiated")
population$AddRecruitment(recruitment$id, 1, area1$id)
population$SetInitialDeviations(initial_deviations$id, area1$id, "undifferentiated")
population$SetGrowth(growth$id)
population$sex_ratio <- 0.5
```

### Fleet settings
```{r}
# Catch index values and observation errors
catch_index <- new(r4mas$IndexData)
catch_index$values <- em_input$L.obs$fleet1
catch_index$error <- rep(em_input$cv.L$fleet1, times = om_input$nyr)
# Catch composition data
catch_comp <- new(r4mas$AgeCompData)
catch_comp$values <- as.vector(t(em_input$L.age.obs$fleet1))
catch_comp$sample_size <- rep(em_input$n.L$fleet1, nyears * nseasons)
# Likelihood component settings
fleet_index_comp_nll <- new(r4mas$Lognormal)
fleet_index_comp_nll$use_bias_correction <- FALSE
fleet_age_comp_nll <- new(r4mas$Multinomial)
# Fleet selectivity settings
fleet_selectivity <- new(r4mas$LogisticSelectivity)
fleet_selectivity$a50$value <- om_input$sel_fleet$fleet1$A50.sel
fleet_selectivity$a50$estimated <- TRUE
fleet_selectivity$a50$phase <- 2
fleet_selectivity$a50$min <- 0.0
fleet_selectivity$a50$max <- max(om_input$ages)
fleet_selectivity$slope$value <- 1 / om_input$sel_fleet$fleet1$slope.sel
fleet_selectivity$slope$estimated <- TRUE
fleet_selectivity$slope$phase <- 2
fleet_selectivity$slope$min <- 0.0001
fleet_selectivity$slope$max <- 5
# Fishing mortality settings
fishing_mortality <- new(r4mas$FishingMortality)
fishing_mortality$estimate <- TRUE
fishing_mortality$phase <- 1
fishing_mortality$min <- 0.0
fishing_mortality$max <- 4
fishing_mortality$SetValues(om_output$f)
# Create the fleet
fleet <- new(r4mas$Fleet)
fleet$AddIndexData(catch_index$id, "undifferentiated")
fleet$AddAgeCompData(catch_comp$id, "undifferentiated")
fleet$SetIndexNllComponent(fleet_index_comp_nll$id)
fleet$SetAgeCompNllComponent(fleet_age_comp_nll$id)
fleet$AddSelectivity(fleet_selectivity$id, 1, area1$id)
fleet$AddFishingMortality(fishing_mortality$id, 1, area1$id)
```

### Survey settings
```{r}
# Survey index values and observation errors
survey_index <- new(r4mas$IndexData)
survey_index$values <- em_input$survey.obs$survey1
survey_index$error <- rep(em_input$cv.survey$survey1, times = om_input$nyr)
# Survey composition
survey_comp <- new(r4mas$AgeCompData)
survey_comp$values <- as.vector(t(em_input$survey.age.obs$survey1))
survey_comp$sample_size <- rep(em_input$n.survey$survey1, times = om_input$nyr)
# Likelihood component settings
survey_index_comp_nll <- new(r4mas$Lognormal)
survey_index_comp_nll$use_bias_correction <- FALSE
survey_age_comp_nll <- new(r4mas$Multinomial)
# Survey selectivity settings
survey_selectivity <- new(r4mas$LogisticSelectivity)
survey_selectivity$a50$value <- om_input$sel_survey$survey1$A50.sel
survey_selectivity$a50$estimated <- TRUE
survey_selectivity$a50$phase <- 2
survey_selectivity$a50$min <- 0
survey_selectivity$a50$max <- max(om_input$ages)
survey_selectivity$slope$value <- 1 / om_input$sel_survey$survey1$slope.sel
survey_selectivity$slope$estimated <- TRUE
survey_selectivity$slope$phase <- 2
survey_selectivity$slope$min <- 0.0001
survey_selectivity$slope$max <- 5
# Create the survey
survey <- new(r4mas$Survey)
survey$AddIndexData(survey_index$id, "undifferentiated")
survey$AddAgeCompData(survey_comp$id, "undifferentiated")
survey$SetIndexNllComponent(survey_index_comp_nll$id)
survey$SetAgeCompNllComponent(survey_age_comp_nll$id)
survey$AddSelectivity(survey_selectivity$id, 1, area1$id)
# Catchability settings
survey$q$value <- em_input$survey_q$survey1
survey$q$min <- 0
survey$q$max <- 10
survey$q$estimated <- TRUE
survey$q$phase <- 1
```

### Build the MAS model
```{r}
mas_model <- new(r4mas$MASModel)
mas_model$nyears <- nyears
mas_model$nseasons <- nseasons
mas_model$nages <- nages
mas_model$extended_plus_group <- max(om_input$ages)
mas_model$ages <- ages
mas_model$catch_season_offset <- 0.0
mas_model$spawning_season_offset <- 0.0
mas_model$survey_season_offset <- 0.0
mas_model$AddPopulation(population$id)
mas_model$AddFleet(fleet$id)
mas_model$AddSurvey(survey$id)
```
## 5. Run `MAS` and check for gradients
### Run `MAS`, save `MAS` outputs, and reset `MAS`
```{r}
# Run MAS
mas_model$Run()
# Write MAS outputs to a json file
write(mas_model$GetOutput(),
  file = file.path(om_path, "mas_output.json")
)
# Reset MAS for next run
mas_model$Reset()
# Import MAS output
mas_output <- jsonlite::read_json(file.path(om_path, "mas_output.json"))
```
### Value and gradient of estimated paramter
```{r}
parameter <- unlist(mas_output$estimated_parameters$parameters)
parameter_table <- as.data.frame(matrix(parameter, ncol = 3, byrow = TRUE))
colnames(parameter_table) <- c(
  "Parameter",
  "Value",
  "Gradient"
)
parameter_table$Value <- round(as.numeric(parameter_table$Value),
  digits = 6
)
parameter_table$Gradient <- round(as.numeric(parameter_table$Gradient),
  digits = 6
)
parameter_table
```

## 6. Compare estimates from the `MAS` with true values from the OM

### Aggregate true values of key variables from the OM
```{r}
om <- list()
om$biomass <- om_output$biomass.mt
om$abundance <- om_output$abundance / 1000
om$ssb <- om_output$SSB
om$recruit <- om_output$N.age[, 1] / 1000
om$f <- apply(om_output$FAA, 1, max)
om$landing <- om_output$L.mt$fleet1
om$survey <- om_output$survey_index$survey1
om$msy <- om_output$msy$msy
om$fmsy <- round(om_output$msy$Fmsy, digits = 3)
om$ssbmsy <- om_output$msy$SSBmsy
om$fratio <- om$f / om$fmsy
om$ssbratio <- om$ssb / om$ssbmsy
om$agecomp <- apply(om_output$N.age / 1000, 1, function(x) x / sum(x))
om$r0 <- om_input$R0 / 1000
om$q <- om_output$survey_q
om$selexparm_fleet <- om_input$sel_fleet
om$selexparm_survey <- om_input$sel_survey
om$recruit_deviation <- om_input$logR.resid
```

### Aggregate estimates of key variables from the `MAS`
Still need to add R0, biological reference points, and selectivity-at-age to `MAS` outputs.
```{r}
popdy <- mas_output$population_dynamics
pop <- popdy$populations[[1]]
flt <- popdy$fleets[[1]]
srvy <- popdy$surveys[[1]]
mas <- list()
mas$biomass <- unlist(pop$undifferentiated$biomass$values)
mas$abundance <- unlist(pop$undifferentiated$abundance$values)
mas$ssb <- unlist(pop$undifferentiated$spawning_stock_biomass$values)
mas$recruit <- unlist(pop$undifferentiated$recruits$values)
mas$f <- unlist(pop$undifferentiated$fishing_mortality$values)
mas$landing <- unlist(flt$undifferentiated$catch_biomass$values)
mas$survey <- unlist(srvy$undifferentiated$survey_biomass$values)
mas$agecomp <- apply(
  matrix(unlist(pop$undifferentiated$numbers_at_age$values),
    nrow = popdy$nyears,
    ncol = popdy$nages,
    byrow = T
  ),
  1,
  function(x) x / sum(x)
)
mas$r0 <- exp(parameter_table$Value[parameter_table$Parameter == "log_R0_1"])
mas$q <- list(parameter_table$Value[parameter_table$Parameter == "q_1"] / 1000)
mas$selexparm_fleet <- list(
  a50 = parameter_table$Value[parameter_table$Parameter == "logistic_selectivity_a50_1"],
  slope = parameter_table$Value[parameter_table$Parameter == "logistic_selectivity_slope_1"]
)
mas$selexparm_survey <- list(
  a50 = parameter_table$Value[parameter_table$Parameter == "logistic_selectivity_a50_2"],
  slope = parameter_table$Value[parameter_table$Parameter == "logistic_selectivity_slope_2"]
)
mas$recruit_deviation <- parameter_table[grep("recruitment_deviations", parameter_table$Parameter), "Value"] # Is the order correct from starting year to ending year?
```

### Generate comparison figures
Compare temporal trends of biomass(B), abundance(A), spawning stock biomass (SSB), recruit (R), fishing mortality (F), Landings (L), and Survey index (SI) from the OM (dots) and `MAS` (lines).
```{r}
par(mfrow = c(4, 2), mar = c(3, 3, 0, 0))
col <- nmfspalette::nmfs_palette("regional web")(2)
var <- c(
  "biomass", "abundance", "ssb", "recruit", "f",
  "landing", "survey"
)
ylab <- c(
  "B (mt)", "A (1000 fish)",
  "SSB (mt)", "R (1000 fish)",
  "F", "L (mt)", "SI (scaled)"
)
for (i in 1:length(var)) {
  ylim <- range(om[[var[i]]], mas[[var[i]]])
  plot(om_input$year, om[[var[i]]],
    xlab = "", ylab = "",
    ylim = ylim, pch = 19,
    col = col[1]
  )
  lines(om_input$year, mas[[var[i]]],
    col = col[2], lty = 1
  )
  mtext("Year", side = 1, line = 2, cex = 0.7)
  mtext(ylab[i], side = 2, line = 2, cex = 0.7)
}
plot.new()
legend("center",
  c("OM", "MAS"),
  pch = c(19, NA),
  lty = c(NA, 1),
  col = col,
  bty = "n"
)
```

Compare age composition from the OM (dots) and `MAS` (lines).
```{r}
par(mfrow = c(8, 4), mar = c(3, 3, 0, 0))
col <- nmfspalette::nmfs_palette("regional web")(2)
var <- c("agecomp")
ylab <- c("Proportion")
for (i in 1:ncol(om[[var]])) {
  ylim <- range(om[[var]][, i], mas[[var]][, i])
  plot(om_input$ages, om[[var]][, i],
    xlab = "", ylab = "",
    ylim = ylim, pch = 19,
    col = col[1]
  )
  lines(om_input$age, mas[[var]][, i],
    col = col[2], lty = 1
  )
  mtext("Age", side = 1, line = 2, cex = 0.7)
  mtext(ylab, side = 2, line = 2, cex = 0.7)
  legend("topright",
    paste("Year", om_input$year[i]),
    bty = "n"
  )
}
plot.new()
legend("topright",
  c("OM", "MAS"),
  pch = c(19, NA),
  lty = c(NA, 1),
  col = col,
  bty = "n"
)
```

Compare recruitment deviations over years from the OM and `MAS`.
```{r}
par(mfrow = c(1, 1), mar = c(1, 4, 1, 1))
col <- nmfspalette::nmfs_palette("regional web")(2)
barplot(rbind(om$recruit_deviation, mas$recruit_deviation),
  beside = T,
  ylab = "Recruitment Deviations",
  col = col
)
box()
legend("topright",
  c("OM", "MAS"),
  col = col,
  pch = c(15, 15),
  bty = "n"
)
```

### Generate comparison table
Compare estimated R0, q, and selectivity parameters. The estimated selectivity slope values from the `MAS` were converted using 1/slope to match the slope from the OM. 
```{r}
# var <- c("R0", "q", "selexparm_fleet", "selexparm_survey")
summary_table <- matrix(c(
  om$r0, mas$r0,
  om$q$survey1, mas$q[[1]],
  om$selexparm_fleet$fleet1$A50.sel1, mas$selexparm_fleet$a50,
  om$selexparm_fleet$fleet1$slope.sel1, 1 / mas$selexparm_fleet$slope,
  om$selexparm_survey$survey1$A50.sel1, mas$selexparm_survey$a50,
  om$selexparm_survey$survey1$slope.sel1, 1 / mas$selexparm_survey$slope
),
ncol = 2, byrow = TRUE
)
colnames(summary_table) <- c("OM", "MAS")
rownames(summary_table) <- c(
  "R0", "q",
  "Fleet selectivity A50",
  "Fleet selectivity slope",
  "Survey selectivity A50",
  "Survey selectivity slope"
)
summary_table
```
