---
title: "ASAP Case Study"
date: "Updated on `r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 10)
```

## Install and library packages
```{r, install, eval= FALSE}

remotes::install_github("nmfs-fish-tools/r4MAS")
install.packages("jsonlite")
remotes::install_github("nmfs-general-modeling-tools/nmfspalette")
remotes::install_github("nmfs-fish-tools/fishsad")
```

```{r, load, eval= FALSE}
library(r4MAS)
library(Rcpp)
library(jsonlite)
library(fishsad)
library(nmfspalette)
```

## Read ASAP input file
```{r}
# Working directory
temp_dir <- tempdir()

# Read ASAP input data
asap_input <- fishsad::asap_simple$input$dat
asap_output <- fishsad::asap_simple$output

```

## Convert ASAP inputs to MAS inputs
```{r}

# Load r4MAS module
r4mas <- Rcpp::Module("rmas", PACKAGE = "r4MAS")

# Find the path of dynamically-loaded file with extension .so on Linux, .dylib on OS X or .dll on Windows
libs_path <- system.file("libs", package = "r4MAS")
dll_name <- paste("r4MAS", .Platform$dynlib.ext, sep = "")
if (.Platform$OS.type == "windows") {
  dll_path <- file.path(libs_path, .Platform$r_arch, dll_name)
} else {
  dll_path <- file.path(libs_path, dll_name)
}
r4mas <- Rcpp::Module("rmas", dyn.load(dll_path))
# General settings
nyears <- asap_input$n_years

nseasons <- 1

nages <- asap_input$n_ages

ages <- 1:asap_input$n_ages

area1 <- new(r4mas$Area)
area1$name <- "area1"

# Recruitment settings
recruitment <- new(r4mas$BevertonHoltRecruitment)

recruitment$R0$value <- asap_input$SR_scalar_ini / 1000
recruitment$R0$estimated <-
  ifelse(asap_input$phase_SR_scalar < 0, FALSE, TRUE) # TRUE
recruitment$R0$phase <- abs(asap_input$phase_SR_scalar)

recruitment$h$value <- asap_input$steepness_ini
recruitment$h$estimated <-
  ifelse(asap_input$phase_steepness < 0, FALSE, TRUE) # TRUE
recruitment$h$phase <- abs(asap_input$phase_steepness)
recruitment$h$min <- 0.2001
recruitment$h$max <- 1.0

recruitment$sigma_r$value <- sqrt(log((asap_input$recruit_cv[1, ])^2 + 1))
recruitment$sigma_r$estimated <- FALSE
recruitment$sigma_r$min <- 0
recruitment$sigma_r$max <- 3
recruitment$sigma_r$phase <- 2

recruitment$estimate_deviations <-
  ifelse(asap_input$phase_rec_devs < 0, FALSE, TRUE) # TRUE
recruitment$constrained_deviations <- TRUE
recruitment$deviations_min <- -15.0
recruitment$deviations_max <- 15.0
recruitment$deviation_phase <- abs(asap_input$phase_rec_devs)
recruitment$SetDeviations(rnorm(nyears, mean = 0, sd = sqrt(log((asap_input$recruit_cv[1, ])^2 + 1))))
recruitment$use_bias_correction <- FALSE

# Growth settings
growth <- new(r4mas$VonBertalanffyModified)

fleet_num <- asap_input$n_fleets
catch_waa_pointer <- asap_input$WAA_pointers[fleet_num * 2 + 1]
catch_empirical_weight <-
  as.vector(t(asap_input$WAA_mats[[catch_waa_pointer]])) # Total catch

ssb_waa_pointer <- asap_input$WAA_pointers[fleet_num * 2 + 2 + 1]
ssb_empirical_weight <-
  as.vector(t(asap_input$WAA_mats[[ssb_waa_pointer]]))

jan1_waa_pointer <- asap_input$WAA_pointers[fleet_num * 2 + 2 + 2]
jan1_empirical_weight <-
  as.vector(t(asap_input$WAA_mats[[jan1_waa_pointer]]))

survey_num <- 1 # Need to be updated using the code below when MAS can have multiple surveys with different unit
# survey_num <- asap_input$n_indices
survey_empirical_weight <- vector(mode = "list", length = survey_num)
for (i in 1:survey_num) {
  survey_waa_pointer <- asap_input$index_WAA_pointers[i]
  survey_waa <- as.vector(t(asap_input$WAA_mats[[survey_waa_pointer]]))

  if (asap_input$index_units[i] == 1) {
    survey_empirical_weight[[i]] <- survey_waa
  } # Survey unit is biomass

  if (asap_input$index_units[i] == 2) {
    survey_empirical_weight[[i]] <- replicate(nages * nyears, 1.0)
  } # Survey unit is number
}

growth$SetUndifferentiatedCatchWeight(catch_empirical_weight)
growth$SetUndifferentiatedWeightAtSeasonStart(jan1_empirical_weight)
growth$SetUndifferentiatedWeightAtSpawning(ssb_empirical_weight)
growth$SetUndifferentiatedSurveyWeight(survey_empirical_weight[[1]])

# Maturity settings
maturity <- new(r4mas$Maturity)
maturity$values <- asap_input$maturity[1, ]

# Natural mortality settings
natural_mortality <- new(r4mas$NaturalMortality)
natural_mortality$SetValues(asap_input$M[1, ])

# Movement settings
movement <- new(r4mas$Movement)
movement$connectivity_females <- c(0.0)
movement$connectivity_males <- c(0.0)
movement$connectivity_recruits <- c(0.0)

# Initial deviations
initial_deviations <- new(r4mas$InitialDeviations)
initial_deviations$values <- rep(0.0, times = nages)
initial_deviations$estimate <-
  ifelse(asap_input$phase_N1_devs < 0, FALSE, TRUE) # TRUE
initial_deviations$phase <- abs(asap_input$phase_N1_devs)

# Create population
population <- new(r4mas$Population)
for (y in 1:(nyears))
{
  population$AddMovement(movement$id, y)
} # y starts from 0 or 1?

population$AddNaturalMortality(natural_mortality$id, area1$id, "undifferentiated")
population$AddMaturity(maturity$id, area1$id, "undifferentiated")
population$AddRecruitment(recruitment$id, 1, area1$id)
population$SetInitialDeviations(initial_deviations$id, area1$id, "undifferentiated")
population$SetGrowth(growth$id)
population$sex_ratio <- 0.5 # need to be updated with sex_ratio <- 1 after resolving the issue here (https://github.com/nmfs-fish-tools/r4MAS/issues/35) to match the assumption from ASAP.

# Catch index values and observation errors
catch_index <- vector(mode = "list", length = fleet_num)
for (i in 1:fleet_num) {
  catch_index[[i]] <- new(r4mas$IndexData)
  catch_index[[i]]$values <- asap_input$CAA_mats[[i]][, (nages + 1)]
  catch_index[[i]]$error <- asap_input$catch_cv[, i]
}

# Catch composition data
catch_comp <- vector(mode = "list", length = fleet_num)
for (i in 1:fleet_num) {
  catch_comp[[i]] <- new(r4mas$AgeCompData)
  catch_comp[[i]]$values <- as.vector(t(asap_input$CAA_mats[[i]][, (1:nages)]))
  catch_comp[[i]]$sample_size <- asap_input$catch_Neff[, i]
}

# Likelihood component settings
fleet_index_comp_nll <- vector(mode = "list", length = fleet_num)
fleet_age_comp_nll <- vector(mode = "list", length = fleet_num)
for (i in 1:fleet_num) {
  fleet_index_comp_nll[[i]] <- new(r4mas$Lognormal)
  fleet_index_comp_nll[[i]]$use_bias_correction <- FALSE

  fleet_age_comp_nll[[i]] <- new(r4mas$Multinomial)
}


# Fleet selectivity settings
fleet_selectivity <- vector(mode = "list", length = fleet_num)
for (i in 1:fleet_num) {
  selectivity_option <- asap_input$sel_block_option[i]

  if (selectivity_option == 1) {
    fleet_selectivity[[i]] <- new(r4mas$AgeBasedSelectivity)
    fleet_selectivity[[i]]$estimated <- TRUE # if it is age based selectivity, can you estimate some values and fix the other values?
    fleet_selectivity[[i]]$phase <- 2 # if it is age based selectivity, can you estimate some values and fix the other values?
    # fleet_selectivity$estimated <-
    #   ifelse(asap_input$sel_ini[[i]][(1:nages), 2] < 0, FALSE, TRUE)
    # fleet_selectivity$phase <- asap_input$sel_ini[[i]][(1:nages), 2]
    fleet_selectivity[[i]]$values <- asap_output$fleet.sel.mats$sel.m.fleet1[1, ]
    # fleet_selectivity[[i]]$values <- asap_input$sel_ini[[i]][(1:nages),1]
  }

  if (selectivity_option == 2) {
    fleet_selectivity[[i]] <- new(r4mas$LogisticSelectivity)
    fleet_selectivity[[i]]$a50$value <- asap_input$sel_ini[[i]][(nages + 2), 1]
    fleet_selectivity[[i]]$a50$estimated <-
      ifelse(asap_input$sel_ini[[i]][(nages + 2), 2] < 0, FALSE, TRUE)
    fleet_selectivity[[i]]$a50$phase <- asap_input$sel_ini[[i]][(nages + 2), 2]
    fleet_selectivity[[i]]$a50$min <- 0.0001
    fleet_selectivity[[i]]$a50$max <- nages

    fleet_selectivity[[i]]$slope$value <- asap_input$sel_ini[[i]][(nages + 1), 1]
    fleet_selectivity[[i]]$slope$estimated <- ifelse(asap_input$sel_ini[[i]][(nages + 1), 2] < 0, FALSE, TRUE)
    fleet_selectivity[[i]]$slope$phase <- asap_input$sel_ini[[i]][(nages + 1), 2]
    fleet_selectivity[[i]]$slope$min <- 0.0001
    fleet_selectivity[[i]]$slope$max <- nages
  }

  # Add double-logistic case later
}

# Fishing mortality settings
fishing_mortality <- new(r4mas$FishingMortality)
fishing_mortality$estimate <- TRUE
fishing_mortality$phase <- asap_input$phase_F1
fishing_mortality$min <- 0.0
fishing_mortality$max <- asap_input$Fmax
fishing_mortality$SetValues(rep(asap_input$F1_ini, nyears))

# Create the fleet
fleet <- vector(mode = "list", length = fleet_num)

for (i in 1:fleet_num) {
  fleet[[i]] <- new(r4mas$Fleet)
  fleet[[i]]$AddIndexData(catch_index[[i]]$id, "undifferentiated")
  fleet[[i]]$AddAgeCompData(catch_comp[[i]]$id, "undifferentiated")
  fleet[[i]]$SetIndexNllComponent(fleet_index_comp_nll[[i]]$id)
  fleet[[i]]$SetAgeCompNllComponent(fleet_age_comp_nll[[i]]$id)
  fleet[[i]]$AddSelectivity(fleet_selectivity[[i]]$id, 1, area1$id)
  fleet[[i]]$AddFishingMortality(fishing_mortality$id, 1, area1$id)
}


# Survey index values and observation errors
survey_index <- vector(mode = "list", length = survey_num)
for (i in 1:survey_num) {
  survey_index[[i]] <- new(r4mas$IndexData)
  survey_index[[i]]$values <- asap_input$IAA_mats[[i]][, 2]
  survey_index[[i]]$error <- asap_input$IAA_mats[[i]][, 3]
}

# Survey composition
survey_comp <- vector(mode = "list", length = survey_num)
for (i in 1:survey_num) {
  survey_comp[[i]] <- new(r4mas$AgeCompData)
  survey_comp[[i]]$values <- as.vector(t(asap_input$IAA_mats[[i]][, 4:(4 + nages - 1)]))
  survey_comp[[i]]$sample_size <- asap_input$IAA_mats[[i]][, (4 + nages)]
  survey_comp[[i]]$missing_values <- 0
}

# Likelihood component settings
survey_index_comp_nll <- vector(mode = "list", length = survey_num)
survey_age_comp_nll <- vector(mode = "list", length = survey_num)
for (i in 1:survey_num) {
  survey_index_comp_nll[[i]] <- new(r4mas$Lognormal)
  survey_index_comp_nll[[i]]$use_bias_correction <- FALSE

  survey_age_comp_nll[[i]] <- new(r4mas$Multinomial)
}

# Survey selectivity settings
survey_selectivity <- vector(mode = "list", length = survey_num)
for (i in 1:survey_num) {
  selectivity_option <- asap_input$index_sel_option[i]

  if (selectivity_option == 1) {
    survey_selectivity[[i]] <- new(r4mas$AgeBasedSelectivity)
    survey_selectivity[[i]]$estimated <- FALSE # If it is age based selectivity, can MAS estimates some values and fixes the rest of values?
    survey_selectivity[[i]]$phase <- 1
    # survey_selectivity[[i]]$estimated <- ifelse(asap_input$index_sel_ini[[i]][(1:nages), 2] < 0, FALSE, TRUE)
    # survey_selectivity[[i]]$phase <- asap_input$index_sel_ini[[i]][(1:nages), 2]
    survey_selectivity[[i]]$values <- asap_input$index_sel_ini[[i]][(1:nages), 1]
  }

  if (selectivity_option == 2) {
    survey_selectivity[[i]] <- new(r4mas$LogisticSelectivity)
    survey_selectivity[[i]]$a50$value <- asap_input$index_sel_ini[[i]][(nages + 2), 1]
    survey_selectivity[[i]]$a50$estimated <- ifelse(asap_input$index_sel_ini[[i]][(nages + 2), 2] < 0, FALSE, TRUE)
    survey_selectivity[[i]]$a50$phase <- asap_input$index_sel_ini[[i]][(nages + 2), 2]
    survey_selectivity[[i]]$a50$min <- 0.0001
    survey_selectivity[[i]]$a50$max <- nages

    survey_selectivity[[i]]$slope$value <- asap_input$index_sel_ini[[i]][(nages + 1), 1]
    survey_selectivity[[i]]$slope$estimated <- ifelse(asap_input$index_sel_ini[[i]][(nages + 1), 2] < 0, FALSE, TRUE)
    survey_selectivity[[i]]$slope$phase <- asap_input$index_sel_ini[[i]][(nages + 1), 2]
    survey_selectivity[[i]]$slope$min <- 0.0001
    survey_selectivity[[i]]$slope$max <- nages
  }

  # Add double-logistic case later
}

# Create the survey
survey <- vector(mode = "list", length = survey_num)
for (i in 1:survey_num) {
  survey[[i]] <- new(r4mas$Survey)

  survey[[i]]$AddIndexData(survey_index[[i]]$id, "undifferentiated")
  survey[[i]]$AddAgeCompData(survey_comp[[i]]$id, "undifferentiated")
  survey[[i]]$SetIndexNllComponent(survey_index_comp_nll[[i]]$id)
  survey[[i]]$SetAgeCompNllComponent(survey_age_comp_nll[[i]]$id)
  survey[[i]]$AddSelectivity(survey_selectivity[[i]]$id, 1, area1$id)

  survey[[i]]$q$value <- asap_input$q_ini[i]
  survey[[i]]$q$min <- 0
  survey[[i]]$q$max <- 10
  survey[[i]]$q$estimated <- ifelse(asap_input$phase_q < 0, FALSE, TRUE)
  survey[[i]]$q$phase <- abs(asap_input$phase_q)
}
```

## Build the MAS model
```{r}

mas_model <- new(r4mas$MASModel)

mas_model$nyears <- nyears
mas_model$nseasons <- nseasons
mas_model$nages <- nages
mas_model$extended_plus_group <- max(ages)
mas_model$ages <- ages
mas_model$catch_season_offset <- 0.0
mas_model$spawning_season_offset <- asap_input$fracyr_spawn
mas_model$survey_season_offset <- (asap_input$index_month[1] - 1) / 12

mas_model$AddPopulation(population$id)

for (i in 1:fleet_num) {
  mas_model$AddFleet(fleet[[i]]$id)
}

for (i in 1:survey_num) {
  mas_model$AddSurvey(survey[[i]]$id)
}
```

## Run `MAS`, save `MAS` outputs, and reset `MAS`
```{r}
# Run MAS
mas_model$Run()

# Write MAS outputs to a json file
write(mas_model$GetOutput(),
  file = file.path(temp_dir, "mas_output.json")
)

# Reset MAS for next run
mas_model$Reset()

# Import MAS output
mas_output <- jsonlite::read_json(file.path(temp_dir, "mas_output.json"))
```

## Aggregate estimates of key variables from the `ASAP`
```{r}
# Read ASAP outputs
asap <- list()

asap$biomass <- asap_output$tot.jan1.B
asap$abundance <- apply(asap_output$N.age, 1, sum)
asap$ssb <- asap_output$SSB
asap$recruit <- asap_output$N.age[, 1]
asap$f <- apply(asap_output$fleet.FAA$FAA.directed.fleet1, 1, max)
asap$landing <- asap_output$catch.pred
asap$survey <- asap_output$index.pred$ind01
asap$agecomp <- apply(asap_output$N.age, 1, function(x) x / sum(x))
asap$r0 <- asap_output$SR.parms$SR.R0
asap$h <- asap_output$SR.parms$SR.steepness
asap$q <- asap_output$q.indices[1]
asap$fleet_selectivity <- asap_output$fleet.sel.mats$sel.m.fleet1[1, ]
asap$survey_selectivity <- asap_output$index.sel[1, ]
asap$year <- asap_output$SR.resids$year
asap$recruit_deviation <- asap_output$SR.resids$logR.dev
# asap$initial_deviation <- c(0, asap_std$value[asap_std$name=="log_N_year1_devs"])
```

## Aggregate estimates of key variables from the `MAS`
```{r}

parameter <- unlist(mas_output$estimated_parameters$parameters)
parameter_table <- as.data.frame(matrix(parameter, ncol = 3, byrow = TRUE))
colnames(parameter_table) <- c(
  "Parameter",
  "Value",
  "Gradient"
)
parameter_table$Value <- round(as.numeric(parameter_table$Value),
  digits = 6
)
parameter_table$Gradient <- round(as.numeric(parameter_table$Gradient),
  digits = 6
)

parameter_table
```

```{r}

popdy <- mas_output$population_dynamics
pop <- popdy$populations[[1]]
flt <- popdy$fleets[[1]]
srvy <- popdy$surveys[[1]]

mas <- list()

mas$biomass <- unlist(pop$undifferentiated$biomass$values)
mas$abundance <- unlist(pop$undifferentiated$abundance$values)
mas$ssb <- unlist(pop$undifferentiated$spawning_stock_biomass$values)
mas$recruit <- unlist(pop$undifferentiated$recruits$values)
mas$f <- unlist(pop$undifferentiated$fishing_mortality$values)
mas$landing <- unlist(flt$undifferentiated$catch_biomass$values)
mas$survey <- unlist(srvy$undifferentiated$survey_biomass$values)
mas$agecomp <- apply(
  matrix(unlist(pop$undifferentiated$numbers_at_age$values),
    nrow = popdy$nyears,
    ncol = popdy$nages,
    byrow = T
  ),
  1,
  function(x) x / sum(x)
)
mas$r0 <- exp(parameter_table$Value[parameter_table$Parameter == "log_R0_1"])
mas$h <- parameter_table$Value[parameter_table$Parameter == "h1"]
mas$q <- list(parameter_table$Value[parameter_table$Parameter == "q_1"])
# mas$fleet_selectivity # Where to find selectivity outputs?
# mas$survey_selectivity # Where to find selectivity outputs?
mas$recruit_deviation <- parameter_table[grep("recruitment_deviations", parameter_table$Parameter), "Value"] # Is the order correct from starting year to ending year?
```

## Generate comparison figures
Compare temporal trends of biomass(B), abundance(A), spawning stock biomass (SSB), recruit (R), fishing mortality (F), Landings (L), and Survey index (SI) from `ASAP` (dots) and `MAS` (lines).
```{r}
years <- as.numeric(rownames(asap_output$N.age))
par(mfrow = c(4, 2), mar = c(3, 3, 0, 0))
col <- nmfspalette::nmfs_palette("regional web")(2)
var <- c(
  "biomass", "abundance", "ssb", "recruit", "f",
  "landing", "survey"
)
ylab <- c(
  "B (mt)", "A (1000 fish)",
  "SSB (mt)", "R (1000 fish)",
  "F", "L (mt)", "SI 1"
)
for (i in 1:length(var)) {
  ylim <- range(asap[[var[i]]], mas[[var[i]]])
  plot(years, asap[[var[i]]],
    xlab = "", ylab = "",
    ylim = ylim, pch = 19,
    col = col[1]
  )
  lines(years, mas[[var[i]]],
    col = col[2], lty = 1
  )
  mtext("Year", side = 1, line = 2, cex = 0.7)
  mtext(ylab[i], side = 2, line = 2, cex = 0.7)
}
plot.new()
legend("center",
  c("ASAP", "MAS"),
  pch = c(19, NA),
  lty = c(NA, 1),
  col = col,
  bty = "n"
)
```

Compare age composition from the `ASAP` (dots) and `MAS` (lines).
```{r}
par(mfrow = c(7, 3), mar = c(3, 3, 0, 0))
col <- nmfspalette::nmfs_palette("regional web")(2)
var <- c("agecomp")
ylab <- c("Proportion")
for (i in 1:ncol(asap[[var]])) {
  ylim <- range(asap[[var]][, i], mas[[var]][, i])
  plot(ages, asap[[var]][, i],
    xlab = "", ylab = "",
    ylim = ylim, pch = 19,
    col = col[1]
  )
  lines(ages, mas[[var]][, i],
    col = col[2], lty = 1
  )
  mtext("Age", side = 1, line = 2, cex = 0.7)
  mtext(ylab, side = 2, line = 2, cex = 0.7)
  legend("topright",
    paste("Year", years[i]),
    bty = "n"
  )
}
plot.new()
legend("topright",
  c("ASAP", "MAS"),
  pch = c(19, NA),
  lty = c(NA, 1),
  col = col,
  bty = "n"
)
```


Compare recruitment deviations over years from the OM and `MAS`.
```{r}
par(mfrow = c(1, 1), mar = c(1, 4, 1, 1))
col <- nmfspalette::nmfs_palette("regional web")(2)
barplot(rbind(asap$recruit_deviation, mas$recruit_deviation),
  beside = T,
  ylab = "Recruitment Deviations",
  col = col
)
box()
legend("topright",
  c("ASAP", "MAS"),
  col = c("gray80", "gray20"),
  pch = c(15, 15),
  bty = "n"
)
```

### Generate comparison table
Compare estimated R0, q, and h.  
```{r}
# var <- c("R0", "h", "q")
summary_table <- matrix(c(
  asap$r0, mas$r0,
  asap$h, mas$h,
  asap$q, mas$q[[1]]
),
ncol = 2, byrow = TRUE
)
colnames(summary_table) <- c("ASAP", "MAS")
rownames(summary_table) <- c("R0", "h", "q")
summary_table
```
