---
title: "SS Case Study"
date: "Updated on `r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 10)
```

## Install and library packages
```{r}
# remotes::install_github("nmfs-fish-tools/r4MAS")
# remotes::install_github("Bai-Li-NOAA/DMAS")

library(r4MAS)
library(Rcpp)
library(jsonlite)
library(DMAS)

```

## Read SS input file
```{r}

# Working directory
temp_dir <- tempdir()

# Read SS input data
ss_input <- DMAS::ss_empirical_waa_input

ss_dat <- ss_input$ss_data
ss_ctl <- ss_input$ss_control

fleet_id <- unique(ss_dat$catch$V3)
survey_id <- unique(ss_dat$CPUE$index)

ss_wtatage <- ss_input$ss_wtatage
ss_starter <- ss_input$ss_starter
ss_forecast <- ss_input$ss_projection

```

## Convert SS inputs to MAS inputs
```{r}

# Load r4MAS module
r4mas <- Rcpp::Module("rmas", PACKAGE="r4MAS")

# General settings
styr <- ss_dat$styr
endyr <- ss_dat$endyr
nyears <- ss_dat$endyr - ss_dat$styr + 1

nseasons <- ss_dat$nseas

ages <-  ss_dat$agebin_vector # Use accumulator age or age group from SS? Include age 0 or not? They are not continous numbers.
nages <- length(ages)
age_id <- as.character(ages)

# ages <-  c(0, ss_dat$Nages)
# nages <- length(ages)

ngenders <- ss_dat$Nsexes

area <- vector(mode="list", length=ss_dat$Nareas)
for (i in 1:ss_dat$Nareas){
  area[[i]] <- new(r4mas$Area)
  area[[i]]$name <- paste("area", i, sep="")
}

# Recruitment settings
if (ss_ctl$SR_function == 3) {
  recruitment <- new(r4mas$BevertonHoltRecruitment)
}
if (ss_ctl$SR_function == 2) {
  recruitment <- new(r4mas$RickerRecruitment)
}

recruitment$R0$value <- exp(ss_ctl$SRparm["SR_LN(R0)", "INIT"])
recruitment$R0$estimated <-
  ifelse (ss_ctl$SRparm["SR_LN(R0)", "PHASE"]<0, FALSE, TRUE)
recruitment$R0$phase <- abs(ss_ctl$SRparm["SR_LN(R0)", "PHASE"])
recruitment$R0$min <- ss_ctl$SRparm["SR_LN(R0)", "LO"]
recruitment$R0$max <- ss_ctl$SRparm["SR_LN(R0)", "HI"]

recruitment$h$value <- ss_ctl$SRparm[2, "INIT"]
recruitment$h$estimated <-
  ifelse(ss_ctl$SRparm[2, "PHASE"] < 0, FALSE, TRUE)
recruitment$h$phase <- abs(ss_ctl$SRparm[2, "PHASE"])
recruitment$h$min <- ss_ctl$SRparm[2, "LO"]
recruitment$h$max <- ss_ctl$SRparm[2, "HI"]

recruitment$sigma_r$value <- exp(ss_ctl$SRparm["SR_sigmaR", "INIT"])
recruitment$sigma_r$estimated <-
  ifelse(ss_ctl$SRparm["SR_sigmaR", "PHASE"] < 0, FALSE, TRUE)
# recruitment$sigma_r$estimated <- TRUE # The estimated sigma_r and recruitment deviations are 0 when estimating sigma_r.
recruitment$sigma_r$min <- ss_ctl$SRparm["SR_sigmaR", "LO"]
recruitment$sigma_r$max <- ss_ctl$SRparm["SR_sigmaR", "HI"]
recruitment$sigma_r$phase <- abs(ss_ctl$SRparm["SR_sigmaR", "PHASE"])

recruitment$estimate_deviations <-
  ifelse(ss_ctl$recdev_phase < 0, FALSE, TRUE)
recruitment$constrained_deviations <- TRUE
recruitment$deviations_min <- -15.0
recruitment$deviations_max <- 15.0
recruitment$deviation_phase <- abs(ss_ctl$recdev_phase)
recruitment$SetDeviations(rep(0.1, times=nyears))

# Growth settings
fleet_num <- length(fleet_id)
catch_waa <- vector(mode="list", length=fleet_num)
for (i in 1:fleet_num){

  catch_waa[[i]] <- vector(mode="list", length=ngenders)

  for (j in 1:ngenders){
    waa <-  as.vector(t(
      ss_wtatage[(ss_wtatage$Fleet==fleet_id[i] & ss_wtatage$Sex == j), age_id]
    ))

    if (ss_dat$CPUEinfo[ss_dat$CPUEinfo$Fleet==fleet_id[i], "Units"]==0) {
      catch_waa[[i]][[j]] <- rep(1.0, nages*nyears)
    } # Unit is number

    if (ss_dat$CPUEinfo[ss_dat$CPUEinfo$Fleet==fleet_id[i], "Units"]==1) {
      catch_waa[[i]][[j]] <- waa
    } # Unit is biomass
  }
}

ssb_waa <- jan1_waa <- vector(mode="list", length=ngenders)
for (j in 1:ngenders){
  ssb_waa[[j]] <- as.vector(t(
    ss_wtatage[(ss_wtatage$Fleet==-2 & ss_wtatage$Sex == j), age_id]
  ))

  jan1_waa[[j]] <- as.vector(t(
    ss_wtatage[(ss_wtatage$Fleet==0 & ss_wtatage$Sex == j), age_id]
  ))
}

survey_num <- 1 # How to include multiple surveys
# survey_num <- length(survey_id)
survey_waa <- vector(mode="list", length=survey_num)
for (i in 1:survey_num) {

  survey_waa[[i]] <- vector(mode="list", length=ngenders)

  for (j in 1:ngenders){

    waa <- as.vector(t(
      ss_wtatage[(ss_wtatage$Fleet==survey_id[i] & ss_wtatage$Sex == j), age_id]
    ))

    if (ss_dat$CPUEinfo[ss_dat$CPUEinfo$Fleet==survey_id[i], "Units"]==0) {
      survey_waa[[i]][[j]] <- rep(1.0, nages*nyears)
    } # Unit is number

    if (ss_dat$CPUEinfo[ss_dat$CPUEinfo$Fleet==survey_id[i], "Units"]==1) {
      survey_waa[[i]][[j]] <- waa
    } # Unit is biomass

  }
}

show(r4mas$VonBertalanffyModified)
growth <- new(r4mas$VonBertalanffyModified)
growth$SetFemaleCatchWeight(catch_waa[[1]][[1]])
growth$SetMaleCatchWeight(catch_waa[[1]][[2]])
growth$SetFemaleWeightAtSeasonStart(jan1_waa[[1]])
growth$SetMaleWeightAtSeasonStart(jan1_waa[[2]])
growth$SetFemaleWeightAtSpawning(ssb_waa[[1]])
growth$SetMaleWeightAtSpawning(ssb_waa[[2]])
growth$SetFemaleSurveyWeight(survey_waa[[1]][[1]])
growth$SetMaleSurveyWeight(survey_waa[[1]][[2]]) # How to set empirical weight-at-age for two surveys? One survey unit is biomass and the other survey unit is number. The data are stored in survey_empirical_weight list.

# Maturity settings
show(r4mas$Maturity)
maturity <- vector(mode="list", length=ngenders)
for (j in 1:ngenders){
  maturity[[j]] <- new(r4mas$Maturity)
  maturity[[j]]$values <- rep(1.0, nages) # SS has length-at-maturity and weight-at-age for fleet -2 uses maturity*fecundity
}


# Natural mortality settings
natural_mortality <- vector(mode="list", length=ngenders)
if (ss_ctl$natM_type==0) {
  for (j in 1:ngenders){
    natural_mortality[[j]] <- new(r4mas$NaturalMortality) # No min and max settings
    if (j==1) {
      natural_mortality[[j]]$estimate <-
        ifelse(ss_ctl$MG_parms["NatM_p_1_Fem_GP_1", "PHASE"] < 0, FALSE, TRUE)
      natural_mortality[[j]]$phase <- abs(ss_ctl$MG_parms["NatM_p_1_Fem_GP_1", "PHASE"])
      natural_mortality[[j]]$values <- ss_ctl$MG_parms["NatM_p_1_Fem_GP_1", "INIT"]
    }

    if (j==2) {
      natural_mortality[[j]]$estimate <-
        ifelse(ss_ctl$MG_parms["NatM_p_1_Mal_GP_1", "PHASE"] < 0, FALSE, TRUE)
      natural_mortality[[j]]$phase <- abs(ss_ctl$MG_parms["NatM_p_1_Mal_GP_1", "PHASE"])
      natural_mortality[[j]]$values <- ss_ctl$MG_parms["NatM_p_1_Mal_GP_1", "INIT"]
    }
  }
}



# Movement settings
movement <- new(r4mas$Movement)
movement$connectivity_females <- c(1.0)
movement$connectivity_males <- c(1.0)
movement$connectivity_recruits <- c(1.0)

# Initial deviations
initial_deviations <- vector(mode="list", length=ngenders)

for (j in 1:ngenders){
  initial_deviations[[j]] <- new(r4mas$InitialDeviations)
  initial_deviations[[j]]$values <- rep(0.1, times=nages)
  initial_deviations[[j]]$estimate <- TRUE # Is it true in SS?
  initial_deviations[[j]]$phase <- 2
}

# Create population
population=new(r4mas$Population)
for (y in 0:(nyears))
{
  population$AddMovement(movement$id, y)
}

population$AddNaturalMortality(natural_mortality[[1]]$id, area[[1]]$id, "females")
population$AddNaturalMortality(natural_mortality[[2]]$id, area[[1]]$id, "males")
population$AddMaturity(maturity[[1]]$id, area[[1]]$id, "females")
population$AddMaturity(maturity[[2]]$id, area[[1]]$id, "males")
population$AddRecruitment(recruitment$id, 1, area[[1]]$id)
population$SetInitialDeviations(initial_deviations[[1]]$id, area[[1]]$id, "females")
population$SetInitialDeviations(initial_deviations[[2]]$id, area[[1]]$id, "males")
population$SetGrowth(growth$id)
population$sex_ratio <- ss_ctl$MG_parms["FracFemale_GP_1", "INIT"] # Did SS use female fraction in this example?

# Catch index values and observation errors
catch_index <- vector(mode="list", length=fleet_num)
for (i in 1:fleet_num){
  catch_index[[i]] <- new(r4mas$IndexData)
  catch_index[[i]]$values <- ss_dat$catch$V4[ss_dat$catch$V3==fleet_id[i] &
                                               ss_dat$catch$V1>0]
  catch_index[[i]]$error <- ss_dat$catch$V5[ss_dat$catch$V3==fleet_id[i] &
                                              ss_dat$catch$V1>0]
}

# Catch composition data
catch_comp <- vector(mode="list", length=fleet_num)
for (i in 1:fleet_num){
  catch_comp[[i]] <- vector(mode="list", length=ngenders)
  for (j in 1:ngenders){
    catch_comp[[i]][[j]] <- new(r4mas$AgeCompData)

    if (j==1){
      catch_comp[[i]][[j]]$values <- as.vector(t(
        ss_dat$agecomp[ss_dat$agecomp$FltSvy==fleet_id[i], 10:(10+nages-1)]
      ))
      catch_comp[[i]][[j]]$sample_size <- ss_dat$agecomp[ss_dat$agecomp$FltSvy==fleet_id[i], "Nsamp"]/2 # Should sample size be divided by 2?
      catch_comp[[i]][[j]]$sex <- "females"
    }

    if (j==2) {
      catch_comp[[i]][[j]]$values <- as.vector(t(
        ss_dat$agecomp[ss_dat$agecomp$FltSvy==fleet_id[i], (10+nages):ncol(ss_dat$agecomp)]
      ))
      catch_comp[[i]][[j]]$sample_size <- ss_dat$agecomp[ss_dat$agecomp$FltSvy==fleet_id[i], "Nsamp"]/2 # Should sample size be divided by 2?
      catch_comp[[i]][[j]]$sex <- "males"
    }
  }
}

# Likelihood component settings
fleet_index_comp_nll <- vector(mode="list", length=fleet_num)
fleet_age_comp_nll <- vector(mode="list", length=fleet_num)
for (i in 1:fleet_num){
  fleet_index_comp_nll[[i]] <- new(r4mas$Lognormal)
  fleet_index_comp_nll[[i]]$use_bias_correction <- FALSE

  fleet_age_comp_nll[[i]] <- new(r4mas$Multinomial)
}


# # Fleet selectivity settings
fleet_selectivity <- vector(mode="list", length=fleet_num)
for (i in 1:fleet_num){
  selectivity_option <- ss_ctl$age_selex_types$Pattern[fleet_id[i]]

  if (selectivity_option==17) {
    fleet_selectivity[[i]] <- new(r4mas$AgeBasedSelectivity)
    fleet_selectivity[[i]]$estimated <- TRUE # if it is age based selectivity, can you estimate some values and fix the other values?
    fleet_selectivity[[i]]$phase <- 2 # if it is age based selectivity, can you estimate some values and fix the other values?
    # fleet_selectivity$estimated <-
    #   ifelse(asap_input$sel_ini[[i]][(1:nages), 2] < 0, FALSE, TRUE)
    # fleet_selectivity$phase <- asap_input$sel_ini[[i]][(1:nages), 2]
    fleet_selectivity[[i]]$values <- seq(0, 1, length.out=nages)
  } # SS uses random walk?

  # Add simple-logistic and double-logistic cases later

}

# Fishing mortality settings
fishing_mortality <- new(r4mas$FishingMortality)
fishing_mortality$estimate <- TRUE
fishing_mortality$phase <- 1
fishing_mortality$min <- 0.0
fishing_mortality$max <- ss_ctl$maxF
fishing_mortality$SetValues(rep(0.01, nyears))

# Create the fleet
fleet <- vector(mode="list", length=fleet_num)

for (i in 1:fleet_num){
  fleet[[i]] <- new(r4mas$Fleet)
  fleet[[i]]$AddIndexData(catch_index[[i]]$id, "undifferentiated")
  fleet[[i]]$AddAgeCompData(catch_comp[[i]][[1]]$id, "males")
  fleet[[i]]$AddAgeCompData(catch_comp[[i]][[2]]$id, "females")
  fleet[[i]]$SetIndexNllComponent(fleet_index_comp_nll[[i]]$id)
  fleet[[i]]$SetAgeCompNllComponent(fleet_age_comp_nll[[i]]$id)
  fleet[[i]]$AddSelectivity(fleet_selectivity[[i]]$id, 1, area[[1]]$id)
  fleet[[i]]$AddFishingMortality(fishing_mortality$id, 1, area[[1]]$id)
}


# Survey index values and observation errors
survey_index <- vector(mode="list", length=survey_num)
for (i in 1:survey_num){
  survey_index[[i]] <- new(r4mas$IndexData)
  survey_index[[i]]$values <- ss_dat$CPUE$obs[ss_dat$CPUE$index==survey_id[i]]
  survey_index[[i]]$error <- ss_dat$CPUE$se_log[ss_dat$CPUE$index==survey_id[i]] # How to deal with missing data?
}

# Survey composition
survey_comp <- vector(mode="list", length=survey_num)
for (i in 1:survey_num){
  survey_comp[[i]] <- vector(mode="list", length=ngenders)
  for (j in 1:ngenders){
    survey_comp[[i]][[j]] <- new(r4mas$AgeCompData)

    if (j==1) {
      survey_comp[[i]][[j]]$values <- as.vector(t(
        ss_dat$agecomp[ss_dat$agecomp$FltSvy==survey_id[i], 10:(10+nages-1)]
      ))
      survey_comp[[i]][[j]]$sample_size <- ss_dat$agecomp[ss_dat$agecomp$FltSvy==survey_id[i], "Nsamp"]/2 # Should sample size be divided by 2?
      survey_comp[[i]][[j]]$sex <- "females"
    }

    if (j==2) {
      survey_comp[[i]][[j]]$values <- as.vector(t(
        ss_dat$agecomp[ss_dat$agecomp$FltSvy==survey_id[i], (10+nages):ncol(ss_dat$agecomp)]
      ))
      survey_comp[[i]][[j]]$sample_size <- ss_dat$agecomp[ss_dat$agecomp$FltSvy==survey_id[i], "Nsamp"]/2 # Should sample size be divided by 2?
      survey_comp[[i]][[j]]$sex <- "males"
    }

  }
}

# Likelihood component settings
survey_index_comp_nll <- vector(mode="list", length=survey_num)
survey_age_comp_nll <- vector(mode="list", length=survey_num)
for (i in 1:survey_num){
  survey_index_comp_nll[[i]] <- new(r4mas$Lognormal)
  survey_index_comp_nll[[i]]$use_bias_correction <- FALSE

  survey_age_comp_nll[[i]] <- new(r4mas$Multinomial)
}

# Survey selectivity settings
survey_selectivity <- vector(mode="list", length=survey_num)
for (i in 1:survey_num){
  selectivity_option <- ss_ctl$age_selex_types$Pattern[survey_id[i]]

  if (selectivity_option==17) {
    survey_selectivity[[i]] <- new(r4mas$AgeBasedSelectivity)
    survey_selectivity[[i]]$estimated <- TRUE # if it is age based selectivity, can you estimate some values and fix the other values?
    survey_selectivity[[i]]$phase <- 2
    # survey_selectivity[[i]]$estimated <- ifelse(asap_input$index_sel_ini[[i]][(1:nages), 2] < 0, FALSE, TRUE)
    # survey_selectivity[[i]]$phase <- asap_input$index_sel_ini[[i]][(1:nages), 2]
    survey_selectivity[[i]]$values <- seq(0, 1, length.out=nages)

  }

  # Add simple-logistic and double-logistic cases later

}

# Create the survey
survey <- vector(mode="list", length=survey_num)
for (i in 1:survey_num){
  survey[[i]] <- new(r4mas$Survey)

  survey[[i]]$AddIndexData(survey_index[[i]]$id, "undifferentiated")
  survey[[i]]$AddAgeCompData(survey_comp[[i]][[1]]$id, "females")
  survey[[i]]$AddAgeCompData(survey_comp[[i]][[2]]$id, "males")
  survey[[i]]$SetIndexNllComponent(survey_index_comp_nll[[i]]$id)
  survey[[i]]$SetAgeCompNllComponent(survey_age_comp_nll[[i]]$id)
  survey[[i]]$AddSelectivity(survey_selectivity[[i]]$id, 1, area[[1]]$id)

  survey[[i]]$q$value <- exp(ss_ctl$Q_parms[paste("LnQ_base_", survey_id[i], "_S", i, sep=""), "INIT"])/1000
  survey[[i]]$q$min <- 0
  survey[[i]]$q$max <- 10
  survey[[i]]$q$estimated <- ifelse(ss_ctl$Q_parms[paste("LnQ_base_", survey_id[i], "_S", i, sep=""), "PHASE"]<0, FALSE, TRUE)
  survey[[i]]$q$phase <- abs(ss_ctl$Q_parms[paste("LnQ_base_", survey_id[i], "_S", i, sep=""), "PHASE"])
}

```

## Build the MAS model
```{r}

mas_model <- new(r4mas$MASModel)

mas_model$nyears <- nyears
mas_model$nseasons <- nseasons
mas_model$nages <- nages
mas_model$extended_plus_group <- nages # what if the age are grouped?
mas_model$ages <- ages
mas_model$catch_season_offset <- 0.0
mas_model$spawning_season_offset <- (ss_dat$spawn_month-1)/12
mas_model$survey_season_offset <- (unique(ss_dat$CPUE$seas[ss_dat$CPUE$index==2])-1)/12

mas_model$AddPopulation(population$id)

for (i in 1:fleet_num){
  mas_model$AddFleet(fleet[[i]]$id)
}

for (i in 1:survey_num){
  mas_model$AddSurvey(survey[[i]]$id)
}

```

## Run `MAS`, save `MAS` outputs, and reset `MAS`
```{r}
# # Run MAS
# mas_model$Run()
# 
# # Write MAS outputs to a json file
# write(mas_model$GetOutput(),
#       file=file.path(data_dir, "mas_output.json"))
# 
# # Reset MAS for next run
# mas_model$Reset()
# 
# # Import MAS output
# mas_output <- jsonlite::read_json(file.path(data_dir, "mas_output.json"))

```

## Questions 
### r4MAS inputs and outputs
- SS uses real ages in weight-at-age data, but age group in age composition data. How to set up ages in r4MAS in this situation? 

- There are missing data in inputs (e.g., survey index and age composition).  How to set up survey in r4MAS with missing data? 
