---
title: "Introduction"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This package allows you to build and run a Metapopulation Assessment System (MAS) stock assessment model directly from R. It also includes functions to inter-translate between different assessment platforms. In this vignette, we detail how to set up a MAS model and run it from R using dummy data included in the package. It is straightforward to replace this with your own data.

The first step is to ensure you have `Rcpp`, the R package that allows calling of C++ code directly from R, and `jsonlite`, the R package that facilitates reading input and output files from JSON format. Right now, the `RMAS` package is only available on Github so you need to install it with the `remotes` package (this package is included in `devtools` if you have already installed that.)

```{r setup_pkgs, eval= FALSE}
##install.packages("Rcpp")
##install.packages("jsonlite")
##install.packages("remotes")

remotes::install_github("nmfs-fish-tools/RMAS")

```


Next, you will need to load the .dll included in the RMAS package.

```{r dll_load, echo = FALSE}
require(Rcpp)
require(RMAS)
d <- system.file(package = "RMAS")
r4mas <-
  Module("rmas", dyn.load(paste(d,"/libs/x64/RMAS", .Platform$dynlib.ext, sep = "")))
```


## Setting model specifications

Next, we define model scalars, such as the number of years of the model, the number of seasons, and the vector of ages. For this example, we will use the test data created by `rmas::write_test_data()`, but this can be replaced by user input data later on in the script.

```{r model_config, echo=FALSE}

nyears<-30
nseasons<-1
ages<-c(0.1, seq(1,11))
nages <- length(ages)


input_data <- write_test_data(nyears = nyears)
```

RMAS uses classes to parallel the object-oriented structure of MAS. In R, model classes are initiralized using the `new()` command. For each model component, the associated class is initialized and then its attributes are populated. To view available attributes for each class, use the `show()` function as shown below. 

```{r}
#View attributes and methods associated with each class
show(r4mas$Area)

#define area
area1<-new(r4mas$Area)
area1$name<-"area1"
```

## Setting parameter values

Next, we populate the parameter constructors for each different set of parameters: recruitment, growth, maturity, mortality, and movement. For functions with several parameters (e.g. the Beverton-Holt recruitment function), the `create_par_section()` function exists. This takes a number of arguments: the type of section, the associated object created for that function, then a series of vectors of length equivalent to the number of parameters. These vectors denote the name, lower bound, upper bound, units, phase, and value for each parameter. If any parameter doesn't have one of these attributes, you can use an NA in the vector. Or, if one of the attributes is not defined for all attributes (e.g. units in this case), you can pass a single NA for that attribute or leave the argument blank.

```{r}
#Recruitment
recruitment<-new(r4mas$BevertonHoltRecruitment)

devs_list <- list(TRUE, TRUE, rep(0.0,22))
recruitment <- create_par_section(section_type = "recruitment", section_type_object = recruitment, par_names = c("R0","h","sigma_r","recdevs"),
                   par_lo = c(NA, 0.2001,NA,-15), par_hi = c(NA,1.0,NA,15), par_units = NA, par_phase = c(1,-2,-1,1), par_value = c(1000,0.75,0.55, NA), 
                   rec_devs=devs_list)


#Initial Deviations
initial_deviations<-new(r4mas$InitialDeviations)
initial_deviations$values<-rep(0.0,nages)
initial_deviations$estimate<-FALSE
initial_deviations$phase<-1


#Growth
growth<-new(r4mas$VonBertalanffyModified)
empirical_weight <- input_data$ewaa
growth$SetUndifferentiatedCatchWeight(empirical_weight)
growth$SetUndifferentiatedSurveyWeight(empirical_weight)
growth$SetUndifferentiatedWeightAtSeasonStart(empirical_weight)
growth$SetUndifferentiatedWeightAtSpawning(empirical_weight)
growth <- create_par_section(section_type="growth", section_type_object = growth, 
                             par_names = c("a_min","a_max","c","lmin","lmax","alpha_f","alpha_m","beta_f","beta_m"),
                             par_value=c(0.01,10.0,0.3,5,50,2.5E-5,2.5E-5,2.9624,2.9624))

#Maturity
maturity<-new(r4mas$Maturity)
maturity$values<-c(0.011488685,
                   0.16041065,
                   0.45232527,
                   0.497389935,
                   0.499869405,
                   0.499993495,
                   0.499999675,
                   0.499999985,
                   0.5,
                   0.5,
                   0.5,
                   0.5)

#Natural Mortality
natural_mortality<-new(r4mas$NaturalMortality)
natural_mortality$SetValues(rep(0.2,nages))
#need to add reset button to static variable so you can run again

#define Movement (only 1 area in this model)
movement<-new(r4mas$Movement)
movement$connectivity_females<-c(1.0)
movement$connectivity_males<-c(1.0)
movement$connectivity_recruits<-c(1.0)

```

Next we define the functions related to fishing, i.e. fishing mortality and selectivity.

```{r fishing}
#Fishing Mortality
fishing_mortality<-new(r4mas$FishingMortality)
fishing_mortality$estimate<-TRUE
fishing_mortality$phase<-1
fishing_mortality$min<-1e-8
fishing_mortality$max<-30
fishing_mortality$SetValues(rep(0.3,30))

#Selectivity Model
fleet_selectivity<-new(r4mas$LogisticSelectivity)
fleet_selectivity <- create_par_section("selectivity",fleet_selectivity, par_names = c("a50","slope"),
                                        par_lo = c(0.0,0.0001), par_hi = c(10.0,5.0), par_phase = c(2,0.5),
                                        par_value = c(1.5,1.5))


survey_selectivity<-new(r4mas$LogisticSelectivity)
survey_selectivity <- create_par_section("selectivity",survey_selectivity, par_names = c("a50","slope"),
                                        par_lo = c(0.0,0.0001), par_hi = c(10.0,5.0), par_phase = c(2,2),
                                        par_value = c(1.0,0.3))


survey2_selectivity<-new(r4mas$AgeBasedSelectivity)
survey2_selectivity$estimated<-FALSE
survey2_selectivity$values<-c(1.0,rep(0.0,nages-1))

```

## Creating the population

Now that the classes for parameter values and movement have been created, you can create a population class and add the relevant classes. 


```{r}
population<-new(r4mas$Population)

for (y in 0:(nyears))
{
  population$AddMovement(movement$id, y)
}

population$AddNaturalMortality(natural_mortality$id,area1$id,"undifferentiated")
population$AddMaturity(maturity$id,area1$id, "undifferentiated")
population$AddRecruitment(recruitment$id,area1$id)
population$SetInitialDeviations(initial_deviations$id, area1$id, "undifferentiated")
population$SetGrowth(growth$id)
```


## Adding data to the model

Next, use the data classes (`IndexData` and `AgeCompData`) to add values, error, and sample sizes for each data set. Indices and error values are the same length as the number of years, `{r} nyears`. Age comp data are passed as a vector of length `nages` x `nyears`. The `sample_size` vectors should be of length `nyears`.


```{r}
#Index data
catch_index<-new(r4mas$IndexData)
catch_index$values <- input_data$catch_index$values
catch_index$error <- input_data$catch_index$error

survey_index<-new(r4mas$IndexData)
survey_index$values <- input_data$survey_index$values
survey_index$error <- input_data$survey_index$error


#Age Comp Data
catch_comp<-new(r4mas$AgeCompData)
catch_comp$values <- input_data$catch_comp$values
catch_comp$sample_size <-input_data$catch_comp$sample_size



survey_comp<-new(r4mas$AgeCompData)
survey_comp$values <- input_data$survey_comp$values
survey_comp$sample_size <-input_data$survey_comp$sample_size
```


## Defining fleets
To make each fishing fleet, you can use the `make_fleet()` function. This takes the fishery data associated with that fleet, the selectivity associated with that fleet, the area, and fishing mortality parameters and returns an `r4mas` fleet object. 

```{r}
#NLL models
fleet_index_comp_nll<-survey_index_comp_nll <- survey2_index_comp_nll <- new(r4mas$Lognormal)
fleet_age_comp_nll<-survey_age_comp_nll <- new(r4mas$MultinomialRobust)


#Fleet
fleet<- make_fleet(r4mas, catch_comp, catch_index, fleet_selectivity, area1, fishing_mortality)

#Survey
survey<-new(r4mas$Survey)
survey$AddAgeCompData(survey_comp$id,"undifferentiated")
survey$AddIndexData(survey_index$id,"undifferentiated")
survey$SetIndexNllComponent(survey_index_comp_nll$id)
survey$SetAgeCompNllComponent(survey_age_comp_nll$id)
survey$AddSelectivity(survey_selectivity$id, 1, area1$id)
survey$q$value<-0.0001
survey$q$min<-0
survey$q$max<-10
survey$q$estimated<-TRUE
survey$q$phase<-1
```

## Build and run the model and write output
Finally, we create a `MASModel` object and add the relevant model configuration parameters, fleets, surveys, and populations to the model object. We call the `Run()` method to run the model and then write the output to a .json file.

```{r}
#build the MAS model
mas_model<-new(r4mas$MASModel)
mas_model$nyears<-nyears
mas_model$nseasons<-nseasons
mas_model$nages<-nages
mas_model$extended_plus_group<-15
mas_model$ages<-ages
mas_model$AddFleet(fleet$id)
mas_model$catch_season_offset<-0.5
mas_model$spawning_season_offset<-0.5
population$sex_ratio<-0.5

mas_model$AddSurvey(survey$id)
mas_model$AddPopulation(population$id)


##########
# Run the model

mas_model$Run()

write(mas_model$GetOutput(), file="mas_s2_output.json")
```

## Read in model output

We use the `jsonlite` package to read in model output. This creates a list in your R environment with the associated output values and estimates. 

```{r}
require(jsonlite)
json_output <- jsonlite::read_json(paste(d,"/extdata/mas_s2_output.json", sep=""))

years <- json_output$nyears
ages <- json_output$nages
popdy <- json_output$population_dynamics

#View the estimated values
#popdy$fleet$undifferentiated
#popdy$survey$undifferentiated



```

